---
title: "Group_01_Analysis.Rmd"
author: "Group_01"
date: "3/14/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(moderndive)
library(skimr)
library(kableExtra)
library(dplyr)
library(readr)
library(Stat2Data)
library(ggplot2)
library(GGally)


## load data from csv files
```{r}
data <- read.csv("dataset1.csv", na.strings = "") %>% rename("Number_of_Family"=7,
                                                             "Income" = 1,
                                                             "FoodExpenditure" = 3,
                                                             "Gender" = 4,
                                                             "Age" = 5,
                                                             "Type" = 6,
                                                             "Area" = 8,
                                                             "HouseAge" = 9,
                                                             "bedrooms" = 10)
glimpse(data)
```

$FoodExpenditure$ is the annual expenditure by the household on food (in Philippine peso)
$Gender$ is the head of the households sex
$Age$ is the head of the households age (in years)
$Type$ is the relationship between the group of people living in the house
$Number_of_Family$ is the number of people living in the house
$Area$ is the floor area of the house (in $m^2$)
$HouseAge$ is the age of the building (in years)
$bedrooms$ is the number of bedrooms in the house
$Electricity$ indicates that if the house have electricity? (1=Yes, 0=No)


```{r}
data$Region <- factor(data$Region)
data$Gender <- factor(data$Gender)
data$Type <- factor(data$Type)    

                                  Income = (Income),
                                  FoodExpenditure = (FoodExpenditure),
                                  Gender = (Gender),
                                  Age = (Age),
                                  Type = (Type),
                                  Area = (Area),
                                  HouseAge = (HouseAge),
                                  bedrooms = (bedrooms),
                                  Electricity = (Electricity))


summary(model)

data_check[c(1:6,8)] %>% ggpairs()
#pchisq(deviance(model),df.residual(model))
```


Check continuous variables
```{r}
continuous <-select_if(data, is.numeric)
summary(continuous)
```

```{r}
ggplot(continuous, aes(x = Age)) + geom_density(alpha = .2, fill = "#FF00DBFF")
top_one_percent <- quantile(data_drop$Age , .99)
top_one_percent 
```

Area

```

HouseAge 
98% is 65, drop the observations above this threshold
```{r}
ggplot(continuous, aes(x = HouseAge)) + geom_density(alpha = .2, fill = "#FF00DBFF")

Check if outliers were all removed
```{r}
summary(data_drop)
```

## Check factor varialbles 

```{r}
factor <- data.frame(select_if(data_drop, is.factor))
ncol(factor)
```

Create graph for each column
```{r}
graph <- lapply(names(factor), 
                function(x) 
                  ggplot(factor, aes(get(x))) +
                  geom_bar(width = 0.1) +
                  theme(axis.text.x = element_text(angle = 90)))
graph
```

## Summary Statistic

visualize the correlation between the variables
```{r}
corr <- data.frame(lapply(data, as.integer)) #Convert data to numeric
ggcorr(corr, method = c("pairwise", "spearman"), 
             nbreaks = 8, 
             hjust = 0.8,
             label = TRUE,
             label_size = 2,
             color = "grey50")
```
## Train/test set

split the data between a train set and a test set 
```{r}
set.seed(1234)
create_train_test <- function(data1, size = 0.8, train = TRUE) {
  n_row = nrow(data1)
  total_row = size * n_row
  train_sample <- 1: total_row
  if (train == TRUE) {
    return (data1[train_sample, ])
  } else {
    return (data1[-train_sample, ])}}

data_train <- create_train_test(data, 0.8, train = TRUE)
data_test <- create_train_test(data, 0.8, train = FALSE)
```

## Generalized Linear Model

The summary of our model reveals interesting information. The performance of a logistic regression is evaluated with specific key metrics.

## Assess the performance of the model

The logistic regression can be evaluated through the output of the glm() function which stored in a list. Below we print the first five elements to see the results.

```{r}

```{r}
accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
accuracy_Test
```



